Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-08 ltx/paper.tex
more writing. new figures.

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. Object apperance parameters include\\\\
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$q_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{5mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliarty variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{5mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

@@

\begin{figure}[h]
        \center{\includegraphics[width=0.5\textwidth]{../img/synth2_alphaimg_hd.png}}
        \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown. Both the sample and data are overlaid on three images from synthetic experiment. The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.  \willie{will show assignments by marker type in addition to color}}
\end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:alpha_overlay_1}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment. \willie{also show pets 2001 result and per-frame overlays for both}.

To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:	inference_to_results}). The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.

% \begin{figure}
% 	\centering
% 	\subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
% 	\subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
% 	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown. \willie{will show assignments by marker type in addition to color}}
% 	\label{fig:pets2001_results}
% \end{figure}

\begin{figure}[h]
        \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
        \caption{\label{fig:alpha_overlay_1} Extracted data and sample from posterior of the model overlaid on three frames from the pets2001 video.}
\end{figure}

\begin{figure}
	\centering
	\subfloat[]{\includegraphics[width=0.6\textwidth]{../img/pets2000_results_1}}
	\subfloat[]{\includegraphics[width=0.40\textwidth]{../img/pets2000_4objects.png}}
	\caption{Extracted data and sample (a.) from posterior of the model overlaid (b.) on a frame from the pets2000 video}
	\label{fig:pets2000_results}
\end{figure}

\begin{figure}
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
  \label{fig:pets2001_imgs}
\end{figure}

\begin{figure}
  \centering             
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a.) PETS2000 and (b.) PETS2001 video datasets.}
  \label{fig:pm_conf}
\end{figure}

@@

\begin{figure}
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
  \label{fig:pets2009_imgs}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=3in]{../img/pets2009_results}
	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown. \willie{will show assignments by marker type in addition to color}}
	\label{fig:pets2009_results}
\end{figure}

@@

% \subsubsection{PETS2004/Caviar Datasets}
% The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.
% \begin{figure}
% 	\centering
% 	\includegraphics[width=3in]{../img/pets2004_results_1}
% 	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlaid on frame $f=20$.}
% 	\label{fig:pets2009_results}
% \end{figure}

\subsection{New Video Datasets}

The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.

\subsubsection{Ants Video?}

\subsubsection{Cells Video?}

>>>>

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. Object apperance parameters include\\\\
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$q_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{5mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{5mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

@@

\begin{figure}[h]
        \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
        \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown. Both the sample and data are overlaid on three images from synthetic experiment. The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.  \willie{will show assignments by marker type in addition to color}}
\end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment. \willie{also show pets 2001 result and per-frame overlays for both}.

\begin{figure}[h]
        \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
        \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
\end{figure}

\begin{figure}[h]
	\centering
	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
	\label{fig:pets2000_results}
\end{figure}

To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}). The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
  \label{fig:pm_conf}
\end{figure}

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames} \willie{these are old images!... haven't replaced yet}}
  \label{fig:pets2001_imgs}
\end{figure}

@@

@@

SMC inference on the PETS2009 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$. The performance metric measures, SFDA and ATA, were computed for each combination of these two parameters. This sensitivity investigation focused on these parameters due to their potential to have a large effect on object detection accuracies. The $\alpha$ values tested included $\{ 0.01, 0.1, 1, 10, 100 \}$, and the $\rho$ values tested included $\{ 0.7, 0.75, 0.8, 0.85, 0.9 \}$.

The following shows the SFDA sensitivity analysis results. The largest value is in bold.
\begin{center}
\begin{tabular}[c]{l l  l  l  l  l  l}
  & & & $\hspace{8mm}\rho$ & & & \\
  SFDA  &  	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
  &$0.01$ 	& $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
  &$0.1$  	& $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
$\alpha$ 	& $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
  &$10$   	& $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
  &$100$  	& $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
\end{tabular}
\end{center}

The following shows the ATA sensitivity analysis results. The largest value is in bold.
\begin{center}
\begin{tabular}[c]{l l  l  l  l  l  l}
		& & & $\hspace{8mm}\rho$ & & & \\
 	 ATA  & 	 	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
	  &$0.01$ 		& $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
		  &$0.1$ 	& $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
$\alpha$ & $1$  	& $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
  		&$10$   	& $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
 		 &$100$ 	& $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
\end{tabular}
\end{center}

% By inspecting the figures, one can see that there is a point where the SFDA and ATA both achieve a maximal value.

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_sfda.png}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_ata.png}}
  \caption{Surface plot of the sensitivity performance metric tables}
  \label{fig:sens_surf}
\end{figure}

% \begin{figure}
%   \centering             
%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result1.png}}
%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result2.png}}
%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result3.png}}
%   \caption{Surface of the sensitivity performance metric tables}
%   \label{fig:sens_surf}
% \end{figure}

% \begin{figure}
%   \centering             
%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
%   \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}\willie{these are old images!... haven't replaced yet}}
%   \label{fig:pets2009_imgs}
% \end{figure}

% \begin{figure}
% 	\centering
% 	\includegraphics[width=3in]{../img/pets2009_results}
% 	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
% 	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
% 	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown. \willie{will show assignments by marker type in addition to color}}
% 	\label{fig:pets2009_results}
% \end{figure}

% \subsection{New Video Datasets}

% The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.

% \subsubsection{Ants Video?}

% \subsubsection{Cells Video?}
